### 2019-11-10 Mathematical Statatistics/ch_02.rmd
mathematical statistics, ch2
mathematical statistics, ch2

>>>>

### 2.5

Import the General Social Survey data.

gss <- data.table(read.csv(paste0(data.dir, "GSS2002.csv"),
                 header = T))

a.) Create a table and a bar chart of the response to the question about the death penalty.

table(gss$DeathPenalty)

ggplot(gss, aes(DeathPenalty)) +
   geom_bar(aes(fill = ..count..)) +
   labs(title = "Death Penalty Survey Results")

b.) Use the _table_ command and the summary command in R on the gun ownership variable. What additional information does the summary command give that the table does not?

table(gss$OwnGun)

summary(gss$OwnGun)

The summary tells us how many people didn't respond at all to the question.

c.) Create a contingency table displaying the relationship between opinions about the death penalty to that about gun ownership.

with(gss, {
   table(OwnGun, DeathPenalty)
})

d.) What proportion of gun owners favor the death penalty? Does it appear to be different from the proportion among those who do not own guns?

round(prop.table(with(gss, {
   table(OwnGun, DeathPenalty)
})) * 100, 2)

It does seem that gun owners are overwhelmingly in favor of the death penalty compared to those who do not own a gun.

### 2.6

Import the data from the recidivism case study in section 1.4.

Import the General Social Survey data.

recid <- data.table(read.csv(paste0(data.dir, "Recidivism.csv"),
                 header = T))

a.) Create a table and bar chart  of the __Recid__ variable.

```{r echo = T}
table(recid$Recid)
```

```{r echo = T}
ggplot(recid, aes(Recid)) +
   geom_bar(aes(fill = ..count..)) +
   labs(title = "Recidivism")
```

b.) Create a contingency table summarizing the relationship between recidivism (__Recid__) by age (__Age25__).

```{r echo = T}
with(recid, {
   table(Recid, Age25)
})
```

Of those over 25 years of age, what proportion were sent back to prision?

```{r echo = T}
round(prop.table(with(recid, {
   table(Recid, Age25)
})) * 100, 2)
```

c.) Create side-by-side boxplots of the number of days to recidivism grouped by type of violation, and give three comparative statements about the distributions.

```{r echo = T, fig.height=3.5}
ggplot(recid[!is.na(Days)], aes(Type, Days)) +
   geom_boxplot(aes(fill = Type)) +
   labs(title = "Days Until Recidivism by Crime Type")

+ 1.) Technical violations seem to happen quicker after initial release than new crimes.

+ 2.) There are more outliers present in the technical violations, with a cluster of them occuring after 900 days.

+ 3.) The variance is days until recidivism is larger for new crimes. The chances of a technical violation occuring after 450 days drop dramatically.

```{r echo = T}
recid[!is.na(Days), .(Variance = comma(var(Days))), by = Type]
```

d.) Use the quantile command to obtain the quartiles of the number of days to recidivism. Since there are missing values (__NA__) for those released offeneders who had not recidivated, you will need to add the argument __na.rm = T__ to the __quartile__ command to exclude those observations.

```{r echo = T}
quantile(recid$Days, na.rm = T)
```
e.) Create ecdf's of days to recidivism for those under 25 years of age and those 25 years of age or older.

with(recid, {
   recid$U25 <- Age25 == "Under 25"

   plot.ecdf(recid[U25 == T]$Days)
   plot.ecdf(recid[U25 == F]$Days, col="cornflowerblue", add = T)
})

Approximately what proportion in each age group were sent back to prision 400 days after release?

Of those who were sent back to prision,

recid[!is.na(Days), 
      .(Days, O400 = Days >= 400), by = Age25][,
                                               .(Proportion = sum(O400) / .N), by = Age25]

went back after being release 400 days or more.
